---
title: "P8105 HW1"
author: "Sijda Ahmed (sa4007)"
output: github_document
---

```{r}
# load in necessary packages
library(moderndive)
library(nycflights13)
library(ggplot2)
library(tidyverse)

```


## Problem 1

```{r}
# load the early_january_weather dataset
data("early_january_weather")

# check the names of the variables
head(early_january_weather)
variables = names(early_january_weather)
variables

# check the size of the dataset
num_row = nrow(early_january_weather)
num_row
num_col = ncol(early_january_weather)
num_col

# find the mean temperature
mean_temp = mean(early_january_weather$temp)
mean_temp

```

Description of the dataset: The names of the variables in this dataset are `r variables`. The variable origin is the weather station. The variables year, month, day, and hour correspond to the time of recording. The variables temp and dewp correspond to the temperature and dewpoint in F. The variable humid is the relative humidity. The variables wind_dir, wind_speed, and wind_gust correspond to the wind direction (in degrees), speed, and gust (in mph). The variable precip is the amount of precipitation in inches. The variable pressure is the sea lebel pressure in millibars. The variable visib is the visibility in miles. The variable time_hour is the date and hour of the recording in POSIXct. This dataset has `r num_row` rows and `r num_col` columns. The mean temperature is `r mean_temp` degrees Farenheit. 

```{r}
# create a scatterplot
ggplot(early_january_weather, 
       aes(x = time_hour, 
           y = temp, 
           color = humid)) + 
  ggtitle("Temperature and Humidity Throughout Time") +
  xlab("Time/Hour") +
  ylab("Temperature") +
  geom_point()

# save the scattterplot to my project directory
ggsave("plot_temp.pdf")

```

Description of the patterns in this plot: As the hours and days increase, the temperature seems to have an overall increase in temperature. The temperature also fluctuates up and down as the time increases. The humidity is between between 30 and 80 for most hours. The humidity seems to be really high (80 and up) the two or three days leading up to around January 14. 


## Problem 2

```{r}
# create a dataframe
df = 
  tibble(
    samp_numeric = rnorm(10),
    greater_0 = samp_numeric > 0,
    vec_char = c("This", "is", "a", "character", "vector", "with", "random", "words", "in", "it"),
    vec_factor = factor(c("high", "moderate", "low", "high", "moderate", "low", "high", "moderate", "low", "high"))
  )

```

```{r}
# find the mean of the random sample
mean_samp_numeric = df |>
  pull(samp_numeric) |>
  mean()
mean_samp_numeric

# find the mean of the logical vector
mean_greater_0 = df |>
  pull(greater_0) |>
  mean()
mean_greater_0

# find the mean of the character vector
mean_vec_char = df |>
  pull(vec_char) |>
  mean()
mean_vec_char

# find the mean of the factor vector
mean_vec_factor = df |>
  pull(vec_factor) |>
  mean()
mean_vec_factor

```

Taking the mean of the random sample (numerical vector) and logical vector works. The mean of the random sample is `r mean_samp_numeric` and the mean of the logical vector is `r mean_greater_0`. Taking the mean of the character vector and factor vector does not work. 

```{r echo=TRUE, warning=FALSE, message=FALSE, results='hide'}
# Using as.numeric function on the logical, character, and factor variables
as.numeric(df |> pull(greater_0))
as.numeric(df |> pull(vec_char))
as.numeric(df |> pull(vec_factor))

```

The logical and factor variables can be concerted into numerical values. Using as.numeric() on the logical variable replaces TRUE and FALSE with 1's and 0's. Using as.numeric() on the factor variable replaces the three factor levels with 1's, 2's, and 3's. But the character variable cannot be converted into numerical values and results in NA's when as.numeric() is applied it. 

You can only take the mean of numerical and logical variables. It is possible to take the mean of a logical variable without using as.numeric() because TRUE and FALSE are coded as 1's and 0's. It is not possible to take the mean of a character variable with or without using as.numeric() because character variables cannot be converted to numerical values. It is only possible to take the mean of a factor variable after as.numeric() has been applied to it. 


